---
title: "glmnet and randomForest"
author: "Paolo Angelino"
output:
  html_document:
    depth: 2
    highlight: tango
    number_sections: yes
    toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE, message=FALSE, warning=FALSE}
if (!require("tidyverse", quietly = TRUE))
    install.packages("tidyverse")

if (!require("caret", quietly = TRUE))
    BiocManager::install("caret")

if (!require("caTools", quietly = TRUE))
    BiocManager::install("caTools")

# load libraries
library(tidyverse)
library(caret)
library(caTools)
library(SingleCellExperiment)
```

https://www.kaggle.com/code/petrschonbauer/glmnet-and-randomforest-in-r-caret/script

# Data wrangling

## Load data, remove redundat variables

```{r, message=FALSE, warning=FALSE}
# load data
spe <- readRDS(file = "/data/pangelin/HUG/Thibaud/RC_GEOMX/data/GEO_singlecellexperiment.rds")
```

## Summary of data

```{r}
# propotion in response
table(colData(spe)$Response)
prop.table(table(colData(spe)$Response))

# are there any missing values?
any(is.na(colData(spe)$Response))
```

## filter samples

Keep samples for the 4 datasets for which we have TRG

```{r}
spe <- spe[, spe$batch %in% c("GSE150082", "GSE94104", "GSE133057", "GSE93375")]
spe <- spe[, spe$Response %in% c("yes","no")]
dim(spe)
```

## filter features

Keep only significant genes from meta-analysis 

```{r}
CRT.extrR.markers <- readRDS("/data/pangelin/HUG/Thibaud/RC_GEOMX/data/CRT.extrR.markers.rds")
spe <- spe[intersect(CRT.extrR.markers$Symbol,rownames(spe)),]
#keep <- c(grep("XRCC",rownames(spe)),grep("RAD",rownames(spe)))
#spe <- spe[keep,]

dim(spe)
```

# Models

## Create common validation indicies, `trainControl`

Start with dividing predictors (features) and response (class) into separated variables `df_x` and `df_y`.

```{r}
# subset predictors variables
df_x <- t(assay(spe,2))

# subset response variable
df_y <- spe$Response

# convert diagnosis as factor
df_y <- as.factor(df_y)
```

In order to achieve fair comparison of models we have to train and test models on the same train/test splits. `createFolds` seems to be covinient way how to achieve such splits. The outcome (indices) are used later as an `index` parameter of common `trainControl` object for all tested models.

```{r}
# create indices for each fold
set.seed(1111)
my_folds <- createFolds(df_y, k = 5)
```

`my_folds` now contains indices of held-out (validation) samples of each fold. In addition distribution of class (benign / malignant) is preserved.

```{r}
# structure of "my_fold"
my_folds %>% glimpse
# distribution of class is preserved in each fold
my_folds %>% 
  map_df(~prop.table(table(df_y[.])))
```

```{r}
# create trainControl object
my_trainControl <- trainControl(
  summaryFunction = twoClassSummary,
  classProbs = TRUE,
  verboseIter = FALSE,
  savePredictions = TRUE,
  index = my_folds)
```

## glmnet model

```{r, message=FALSE, warning=FALSE}
# train glmnet model
set.seed(9999)
model_glmnet <- train(x = df_x, y = df_y, 
                   method = "glmnet",
                   metric = "ROC",
                   tuneLength = 20,
                   trControl = my_trainControl, 
                   family = "binomial"
)

# print model
model_glmnet
```

Some remarks:

*  ROC = area under curve
*  Sens = sensitivity = true positive rate = proportion of positives that are correctly identified as such
*  Spec = specificity = true negative rate = proportion of negatives that are correctly identified as such

Another approach to find the model with the highest ROC:

```{r}
model_glmnet$results %>% 
  as_tibble %>%  
  arrange(desc(ROC))
```

```{r}
# plot ROCs
plot(model_glmnet)
```

```{r}
library(plotROC)
# Select a parameter setting
selectedIndices <- model_glmnet$pred$alpha == model_glmnet$bestTune$alpha & model_glmnet$pred$lambda == model_glmnet$bestTune$lambda

g <- ggplot(model_glmnet$pred[selectedIndices, ], aes(m=yes, d=factor(obs, levels = c("no", "yes")))) + 
  geom_roc(n.cuts=0) + 
  coord_equal() +
  style_roc()

g + annotate("text", x=0.75, y=0.25, label=paste("AUC =", round((calc_auc(g))$AUC, 4)))
```


## randomForest model

```{r, message=FALSE, warning=FALSE}
# train randomForest model
model_randomForest <- train(x = df_x, y = df_y, 
                            method = "ranger",
                            metric = "ROC",
                            importance = "permutation",
                            tuneLength = 20,
                            trControl = my_trainControl)

# print model
model_randomForest

# plot model
plot(model_randomForest)
```

```{r}
model_randomForest$results %>% 
  as_tibble %>%  
  arrange(desc(ROC))
```

Now we can compare top 10 importance variables:

```{r}
plot(varImp(model_glmnet, scale = FALSE), top = 10, main = "glmnet")
plot(varImp(model_randomForest, scale = FALSE), top = 10, main = "randomForest")
```

Variables quite differs.

## Compare glmnet and randomForest performance

```{r}
model_list <- list(glmnet = model_glmnet, randomForest = model_randomForest)
resamples <- resamples(model_list)

# plot the comparison
bwplot(resamples, metric = "ROC")
# plot the comparison per each fold
xyplot(resamples, metric = "ROC")
```

# Conclusion

*  both models lead to low ROC values
*  open points: difference between variable importance of both models

