---
title: "Apply models to GEOmx data"
author: "Paolo Angelino"
format: html
editor: visual
---


Load data

```{r}
library(readr)
library(dplyr)
GEOMXcohort_pseudobulk_nuclei <- read_csv2("/data/pangelin/HUG/Thibaud/RC_GEOMX/data/GEOMXcohort_pseudobulk_nuclei.csv")
```

```{r}
test_data <- GEOMXcohort_pseudobulk_nuclei %>% filter(Response != "Mid responders") %>% select(-one_of(c("...1", "Code", "Response"))) 

response <- GEOMXcohort_pseudobulk_nuclei %>% filter(Response != "Mid responders") %>% pull(Response)
response <- ifelse(response == "Good responders", "yes", "no")
```

Load model

```{r}
#| echo: false
# models <- readRDS("/data/pangelin/HUG/Thibaud/RC_GEOMX/analysis/results/randomSampling_modeling_extra_final_de_intersect_plus_bulk_genes_6datasets_scaled.rds")
models <- readRDS("/data/pangelin/HUG/Thibaud/RC_GEOMX/analysis/results/randomSampling_modeling_extra_final_DEintersect_6datasets_scaled.rds")
dim(models$gl$trainingData)
```

```{r}
length(intersect(colnames(models$gl$trainingData), colnames(test_data)))
missing_genes <- setdiff(colnames(models$gl$trainingData), colnames(test_data))
missing_genes <- missing_genes[missing_genes != ".outcome"]
for (n in missing_genes) {
  test_data[,n] <- 0
}
summary(test_data[,1:10])
test_data <- t(scale(t(as.matrix(test_data))))
test_data <- test_data[,intersect(colnames(models$gl$trainingData), colnames(test_data))]
summary(test_data[,1:10])
#
```


## Test


```{r}
  test_res <- predict(models$gl, newdata =  test_data, type="prob")
  print(pROC::plot.roc(response, test_res$yes, print.auc=TRUE))


  test_res <- predict(models$rf, newdata =  test_data, type="prob")
  print(pROC::plot.roc(response, test_res$yes, print.auc=TRUE))

```

# GSE87211

```{r}
# Version info: R 4.2.2, Biobase 2.58.0, GEOquery 2.66.0, limma 3.54.0
################################################################
#   Data plots for selected GEO samples
library(GEOquery)
library(limma)
library(umap)

# load series and platform data from GEO

gset <- getGEO("GSE87211", GSEMatrix =TRUE, getGPL=TRUE, AnnotGPL=TRUE)
if (length(gset) > 1) idx <- grep("GPL13497", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]

ex <- exprs(gset)
# log2 transform
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
LogC <- (qx[5] > 100) ||
          (qx[6]-qx[1] > 50 && qx[2] > 0)
if (LogC) { ex[which(ex <= 0)] <- NaN
  ex <- log2(ex) }

# box-and-whisker plot
dev.new(width=3+ncol(gset)/6, height=5)
par(mar=c(7,4,2,1))
title <- paste ("GSE87211", "/", annotation(gset), sep ="")
boxplot(ex, boxwex=0.7, notch=T, main=title, outline=FALSE, las=2)
dev.off()

# expression value distribution plot
par(mar=c(4,4,2,1))
title <- paste ("GSE87211", "/", annotation(gset), " value distribution", sep ="")
plotDensities(ex, main=title, legend=F)

# mean-variance trend
ex <- na.omit(ex) # eliminate rows with NAs
plotSA(lmFit(ex), main="Mean variance trend, GSE87211")

# UMAP plot (multi-dimensional scaling)
ex <- ex[!duplicated(ex), ]  # remove duplicates
ump <- umap(t(ex), n_neighbors = 15, random_state = 123)
plot(ump$layout, main="UMAP plot, nbrs=15", xlab="", ylab="", pch=20, cex=1.5)
library("maptools")  # point labels without overlaps
pointLabel(ump$layout, labels = rownames(ump$layout), method="SANN", cex=0.6)

gset.bkp <- gset
```

### Filter by treatment

```{r}
dim(gset)
table(gset$`preoperative radiochemotherapy (rct):ch1`)
```

```{r}
gset <- gset[,gset$`preoperative radiochemotherapy (rct):ch1` != "5-FU + RT"]
gset <- gset[,gset$`tissue:ch1` == "rectal tumor"]
gset <- gset[,gset$`depth of invasion before rct:ch1` != "NA" & gset$`depth of invasion after rct:ch1` != "NA"]
```

```{r}
dim(gset)
table(gset$`preoperative radiochemotherapy (rct):ch1`)
```

```{r}
table(gset$`depth of invasion before rct:ch1`, gset$`depth of invasion after rct:ch1`)
```

```{r}
table(gset$`lymph node metastasis before rct:ch1`, gset$`lymph node metastasis after rct:ch1`)
```

```{r}
table(gset$`depth of invasion before rct:ch1`, as.numeric(gset$`depth of invasion before rct:ch1`) - as.numeric(gset$`depth of invasion after rct:ch1`) )
```

```{r}
table(as.numeric(gset$`lymph node metastasis before rct:ch1`) - as.numeric(gset$`lymph node metastasis after rct:ch1`), 
      as.numeric(gset$`depth of invasion before rct:ch1`) - as.numeric(gset$`depth of invasion after rct:ch1`) )
```

```{r}
table(as.numeric(gset$`lymph node metastasis after rct:ch1`), 
      as.numeric(gset$`depth of invasion before rct:ch1`) - as.numeric(gset$`depth of invasion after rct:ch1`) )
```

## define response

```{r}
# remove mid responders
gset <- gset[,as.numeric(gset$`depth of invasion before rct:ch1`) - as.numeric(gset$`depth of invasion after rct:ch1`) != 2]
gset <- gset[,as.numeric(gset$`depth of invasion before rct:ch1`) - as.numeric(gset$`depth of invasion after rct:ch1`) != 1]
dim(gset)
response <- ifelse( as.numeric(gset$`depth of invasion before rct:ch1`) - as.numeric(gset$`depth of invasion after rct:ch1`) > 1, "yes", "no" )
table(response)
```

```{r}
gset <- gset[!duplicated(fData(gset)$GENE_SYMBOL),]
gset <- gset[fData(gset)$GENE_SYMBOL %in% intersect(colnames(models$gl$trainingData), fData(gset)$GENE_SYMBOL),]
dim(gset)
length(intersect(colnames(models$gl$trainingData), fData(gset)$GENE_SYMBOL))
missing_genes <- setdiff(colnames(models$gl$trainingData), fData(gset)$GENE_SYMBOL)
missing_genes <- missing_genes[missing_genes != ".outcome"]

all_genes  <- c(rownames(gset) , missing_genes)
all_genes
```

```{r}
test_data <- ex
test_data <- as.data.frame(test_data)
for (n in missing_genes) {
  test_data[n,] <- 0
}
summary(test_data[,1:10])
test_data <- scale(as.matrix(test_data))
# test_data <- test_data[,intersect(colnames(models$gl$trainingData), colnames(test_data))]
summary(test_data[,1:10])
test_data <- test_data[all_genes,colnames(gset)]
rownames(test_data) <- c(fData(gset)$GENE_SYMBOL, missing_genes)
```


## Test


```{r}
  test_res <- predict(models$gl, newdata =  t(test_data), type="prob")
  print(pROC::plot.roc(response, test_res$yes, print.auc=TRUE))


  test_res <- predict(models$rf, newdata =  t(test_data), type="prob")
  print(pROC::plot.roc(response, test_res$yes, print.auc=TRUE))

```

### Filter by treatment

For our study we selected samples from tumours treated with 5FU infusion alone and a total irradiation dose of 50.4 Gy (28 x 1.8 Gy) where T and N stage before RT was not missing (Suppl Fig 1C). Tumours showing T0N0 stage after RT treatment were classified as pCRs, otherwise as non-pCRs.  

```{r}
gset <- gset.bkp
```

```{r}
#gset <- gset[,gset$`preoperative radiochemotherapy (rct):ch1` == "5-FU + RT"]
gset <- gset[,gset$`tissue:ch1` == "rectal tumor"]
gset <- gset[,gset$`depth of invasion before rct:ch1` != "NA" & gset$`depth of invasion after rct:ch1` != "NA"]
```

```{r}
dim(gset)
```

```{r}
table(gset$`depth of invasion after rct:ch1`, gset$`lymph node metastasis after rct:ch1`)
```

## define response

```{r}
# remove mid responders
gset <- gset[, as.numeric(gset$`depth of invasion after rct:ch1`) != 2]
#gset <- gset[, as.numeric(gset$`depth of invasion after rct:ch1`) != 1]
dim(gset)
response <- ifelse(gset$`depth of invasion after rct:ch1` < 2 & gset$`lymph node metastasis after rct:ch1` == 0, "yes", "no" )
table(response)
```

```{r}
gset <- gset[!duplicated(fData(gset)$GENE_SYMBOL),]
gset <- gset[fData(gset)$GENE_SYMBOL %in% intersect(colnames(models$gl$trainingData), fData(gset)$GENE_SYMBOL),]
dim(gset)
length(intersect(colnames(models$gl$trainingData), fData(gset)$GENE_SYMBOL))
missing_genes <- setdiff(colnames(models$gl$trainingData), fData(gset)$GENE_SYMBOL)
missing_genes <- missing_genes[missing_genes != ".outcome"]

all_genes  <- c(rownames(gset) , missing_genes)
all_genes
```

```{r}
test_data <- ex
test_data <- as.data.frame(test_data)
for (n in missing_genes) {
  test_data[n,] <- 0
}
summary(test_data[,1:10])
test_data <- scale(as.matrix(test_data))
# test_data <- test_data[,intersect(colnames(models$gl$trainingData), colnames(test_data))]
summary(test_data[,1:10])
test_data <- test_data[all_genes,colnames(gset)]
rownames(test_data) <- c(fData(gset)$GENE_SYMBOL, missing_genes)
```


## Test


```{r}
  test_res <- predict(models$gl, newdata =  t(test_data), type="prob")
  print(pROC::plot.roc(response, test_res$yes, print.auc=TRUE))


  test_res <- predict(models$rf, newdata =  t(test_data), type="prob")
  print(pROC::plot.roc(response, test_res$yes, print.auc=TRUE))

```

# Compare gene lists 

```{r}
eBiomed_33genes <- read_csv("/data/pangelin/HUG/Thibaud/RC_GEOMX/data/eBiomed_33genes.csv")
keep <- readRDS("/data/pangelin/HUG/Thibaud/RC_GEOMX/analysis/results/gene_lists.rds")
```

```{r}
for (n in names(keep)) {
  print(n)  
  print(intersect(keep[[n]], eBiomed_33genes$Gene.Symbol))
}
```




